Tsbrowse: an interactive browser for Ancestral Recombination Graphs

### 2025-04-22 manuscript/cover_letter.tex
Update manuscript/cover_letter.tex
Co-authored-by: Jerome Kelleher <jk@well.ox.ac.uk>

    Visual inspection is a crucial part of data analysis pipelines. For omics,
    this
    vital
    infrastructure is provided by stat-of-art genome browsers. The lack of
    tools
    for the
    visual inspection of ARGs is a major gap in a rapidly advancing field. With
    tsbrowse,
    we aim to bridge this gap by presenting interactive views of the nodes,
    edges
    and mutations
    in an ARG. The manuscript describes the applications and architecture of
    tsbrowse, and
    demonstrates user-interactivity and scalability to millions of samples.

>>>>

    Visual inspection is a crucial part of data analysis pipelines. For omics,
    this
    vital
    infrastructure is provided by state-of-art genome browsers. The lack of
    tools
    for the
    visual inspection of ARGs is a major gap in a rapidly advancing field. With
    tsbrowse,
    we bridge this gap by presenting interactive views of the nodes,
    edges
    and mutations
    in an ARG. The manuscript describes the applications and architecture of
    tsbrowse, and
    demonstrates user-interactivity and scalability to millions of samples.

### 2025-04-22 manuscript/paper.tex

\journaltitle{Bioinformatics}
\journaltitle{Preprint}
\DOI{DOI HERE}
\copyrightyear{2024}
% \pubyear{2019}
% \access{Advance Access Publication Date: Day Month Year}
\appnotes{Paper}

@@

\title[tsbrowse]{Tsbrowse: a genome browser for Ancestral Recombination Graphs}

@@

\keywords{Ancestral Recombination Graphs, Holoviz, biobank-scale ARG
    visualisation, genome browser} \maketitle

@@

\section*{Funding}
This work was supported by funding from the Biotechnology and Biological
Sciences Research Council (UKRI-BBSRC) [grant number BB/T008784/1],
\textcolor{red}{add Novo funding}. DMR is funded by 
a studentship from the Wellcome Programme in Genomic Medicine and Statistics.
JK acknowledges EPSRC (research grant EP/X024881/1),
NIH (research grants HG011395 and HG012473)
and the Robertson Foundation.

@@

\section{Supplementary Material}
\subsection{Simulation of truth dataset (Fig. \ref{fig:Figure_2})}
\label{subsec:sweep_simulation}
Ancestral
histories of 300 samples were simulated with \texttt{SweepGenicSelection}
function in \texttt{msprime (version 1.3.3)}. A combination of models was used:
in the recent past, a selective sweep was simulated with a beneficial allele
situated in the middle of a 5 Mb sequence. The frequency of the allele in the
population was set at 0.0001 at the beginning of the sweep. The allele fixed in
the population at a frequency of 0.9999. The strength of selection was set
using the selection coefficient, \textit{s = 0.25}. A time increment,
\textit{dt = 1e-6} was used to step through the sweep. Mutations were added to
the ARG at a rate of 1e-6. For simulating history before occurrence of the
sweep, a standard coalescent model (Hudson's algorithm \citep{Hudson1983}) was
used until coalescence was achieved.

@@

\clearpage
\section{Supplementary Figures}
\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{figures/SuppFig1.png}
    \caption{\textbf{\texttt{tsbrowse} applied to inference methods}
    A screenshot of \texttt{tsbrowse}'s Edges view for tsinfer+tsdate (A),
ARG-Needle (B), Relate (C) and SINGER (D) inferences of the truth dataset
simulated under a selective sweep model (in Figure 2 of the main text). For
SINGER, an examplar ARG from the set of output posterior ARG samples is shown.
The X coordinate represents genomic position, each horizontal segment on the
plot shows the genomic coordinates that the edge spans, and Y coordinate shows
time of either the parent or child node in the edge.}
    \label{fig:Supplementary_Figure_1}
\end{figure}

>>>>

\journaltitle{Bioinformatics}
\journaltitle{Preprint}
\DOI{DOI HERE}
\copyrightyear{2025}
% \pubyear{2019}
% \access{Advance Access Publication Date: Day Month Year}
\appnotes{Paper}

@@

\title[tsbrowse]{Tsbrowse: an interactive browser for Ancestral Recombination Graphs}

@@

\keywords{Ancestral Recombination Graph, HoloViz, data visualisation, interactive browser} \maketitle

@@

\section*{Funding}
SK acknowledges full support and funding from Novo Nordisk Research Centre Oxford,
and funding from the Biotechnology and Biological Sciences Research Council 
(UKRI-BBSRC) [grant number BB/T008784/1]. DMR is funded by 
a studentship from the Wellcome Programme in Genomic Medicine and Statistics.
JK acknowledges EPSRC (research grant EP/X024881/1),
NIH (research grants HG011395 and HG012473)
and the Robertson Foundation.

@@

\section{Supplementary Material}
\subsection{Simulation of truth dataset (Fig. \ref{fig:Figure_2})}
\label{subsec:sweep_simulation}
Ancestral
histories of 300 samples were simulated with \texttt{SweepGenicSelection}
function in \texttt{msprime (version 1.3.3)}. A combination of models was used:
in the recent past, a selective sweep was simulated with a beneficial allele
situated in the middle of a 5 Mb sequence. The frequency of the allele in the
population was set at 0.0001 at the beginning of the sweep. The allele fixed in
the population at a frequency of 0.9999. The strength of selection was set
using the selection coefficient, \textit{s = 0.25}. A time increment,
\textit{dt = 1e-6} was used to step through the sweep. Mutations were added to
the ARG at a rate of 1e-8 per base pair per generation. A recombination rate of 1e-8 per base pair per generation. For simulating history before occurrence of the
sweep, a standard coalescent model (Hudson's algorithm \citep{Hudson1983}) was
used until coalescence was achieved.

@@

\clearpage
\section{Supplementary Figures}
\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{figures/SuppFig1.png}
    \caption{\textbf{\texttt{tsbrowse} applied to inference methods}
    A screenshot of \texttt{tsbrowse}'s Edges view for tsinfer+tsdate,
ARG-Needle, Relate and SINGER inferences of the truth dataset
simulated under a selective sweep model (in Figure 2 of the main text). For
SINGER, an examplar ARG from the set of output posterior ARG samples is shown.
The X coordinate represents genomic position, each horizontal segment on the
plot shows the genomic coordinates that the edge spans, and Y coordinate shows
time of either the parent or child node in the edge.}
    \label{fig:Supplementary_Figure_1}
\end{figure}
